---
title: "R Notebook"
output: html_notebook
---

This first, quite messy, chunk of code reads in the html from hundreds of transcripts and extracts the text and the date of each show. 

```{r}
library(dplyr)
library(rvest)
library(stringr)
begin <- 'http://www.nbcnews.com/id/'
end <- '/ns/meet_the_press-resources/#.WnJ2aahKuUn'
year_urls <- c('8987534','45753152','42784428','40871902','34643492',
               '28452155','23017575','16456474','10558539','9439083')

year_shows <- list()
for(i in seq(1,length(year_urls))){
  year_html <- read_html(paste0(begin,year_urls[i],end))
  poss_urls <- year_html%>%
    html_nodes('span a')
  year_shows[[i]] <- poss_urls%>%
    html_attr('href')%>%
    .[grepl('nbcnews.com',.)]
}
show_urls <- unlist(year_shows)
transcript_urls <- show_urls[grepl('transcript',show_urls)]
good_html <- list()
for(i in seq(1,length(transcript_urls))){
  good_html[[i]] <- try(read_html(transcript_urls[i]))
}
transcript_urls[2]
capture_text <- function(obj){
  show_text <- obj%>%
    html_nodes('div p')%>%
    html_text()%>%
    trimws()
  return(show_text)
}
  
capture_date <- function(obj, text_list){
  show_date <- obj%>%
    html_nodes('div abbr')%>%
    html_text()
  return(rep(show_date, length(text_list)))
}

show_df <- list()
for(i in seq(1,length(good_html))){
  if(length(good_html[[i]]) == 2){
    show_df[[i]] <- data.frame(text = capture_text(good_html[[i]]),
                               date = capture_date(good_html[[i]], 
                                                   capture_text(good_html[[i]])))
  } else {
    print(length(good_html[[i]]))
    show_df[[i]] <- NA
  }
}

all_show_dfs <- bind_rows(show_df[!is.na(show_df)])
```

Next, we split off the speaker from the thing spoken. It's not the cleanest formating because the speaker and their text comes in a single string delimited by a colon. So after extracting speaker, we need to go through and run some diagnostic checks to clean up the list of guests.

```{r}
speaker <- ifelse(grepl(':',all_show_dfs$text),
                    unlist(lapply(strsplit(all_show_dfs$text,
                                           ':', fixed = TRUE), '[', 1)),NA)
shows_speakers <- data.frame(all_show_dfs, data.frame(speaker))%>%
  do(zoo::na.locf(.))


```

```{r}
shows_speakers%>%
  mutate(first = ifelse(substr(speaker, 0,1)==toupper(substr(speaker,0,1)),1,0),
         second = ifelse(substr(speaker, 1,2)==toupper(substr(speaker,1,2)),1,0),
         third = ifelse(substr(speaker, 2,3)==toupper(substr(speaker,2,3)),1,0),
         name_score = first+second+third)%>%
  distinct(speaker,.keep_all = T)%>%
  filter(name_score > 1)%>%
  write.csv('speakers.csv', row.names = F)

read_html('http://www.tvguide.com/tvshows/meet-the-press/cast/203044/')%>%
  html_nodes('div a')%>%
  html_text()%>%
  trimws()
```

speaker <- ifelse(grepl(':',show_text),
                    unlist(lapply(strsplit(show_text,
                                           ':', fixed = TRUE), '[', 1)),NA)
  show_df <- data.frame(show_text, speaker)%>%
    do(zoo::na.locf(.))#%>%mutate(clean_text = str_replace(.$show_text, .$speaker, ''))
  show_df$show_text <- gsub('[[:punct:] ]+',' ',show_df$show_text)
  show_df$show_text <- str_replace(show_df$show_text, show_df$speaker,'')
  show_df$show_text <- str_replace(show_df$show_text, ':','')
  show_df$speaker <- gsub('[[:punct:] ]+',' ',show_df$speaker)
  show_df <- show_df%>%
    mutate(chars = nchar(trimws(show_text)))%>%
    filter(chars>2)%>%
    mutate(utterance = row_number(),
           date = show_date)%>%
    select(-chars)
  return(
    show_df
         )
}